---
title: "Taxonomic data analysis"
output: html_notebook
---

# Taxonomic data analysis

## Loading the preprocessed dataset (normalized)

```{r}
data = read.csv(file="../refined_data/filtered_taxonomic_abundance.csv",
                sep=";",
                stringsAsFactors = TRUE,
                na.strings = "None")
data$ind = as.factor(data$ind)
head(data)
```

## Function to perform statistical test

The aim of the following function is to compute several test to determine if the relative abundance is statistically different between groups.

```{r}
# Getting statistically different samples

diff_test = function(data){

  # Computing test. Significant is a vector of bool that answer the question : is the relative abundance different between control and colistin group.
  significant = c()
  for (class in levels(data$genre)){
    control = subset(data, cat == "Control" & genre == class)$abond
    treated = subset(data, cat == "Colistin" & genre == class)$abond
    if (length(control) >= 3 & length (treated) >= 3){
      shapiro_C = (shapiro.test(control)[[2]] >= 0.05)
      shapiro_T = (shapiro.test(treated)[[2]] >= 0.05)
      if (shapiro_C & shapiro_T){
        fisher = (var.test(control, treated)[[3]] >= 0.05)
        if (fisher){
          difference_test = (t.test(control, treated)[[3]] <= 0.05)
          significant = c(significant, difference_test)
        }else if (length(control)>= 20 & length(treated)>=20){
          difference_test = (t.test(control, treated)[[3]] <= 0.05)
          significant = c(significant, difference_test)
        }else{
          difference_test = (t.test(control, treated, var.equal=F)[[3]] <= 0.05)
          significant = c(significant, difference_test)
        }
      }else if (length(control) >= 30 & length(treated) >= 30){
        difference_test = (t.test(control, treated)[[3]] <= 0.05)
        significant = c(significant, difference_test)
      }else{
        difference_test = (wilcox.test(control, treated)[[3]] <= 0.05)
        significant = c(significant, difference_test)
      }  
    }else{
      significant = c(significant, 2)
    }
  }
  
  # Adding "(*)" to statistically different samples
  data$genre = as.character(data$genre)
  tmp_levels = unique(data$genre)
  for (index in 1:length(significant)){
    print(index)
    if (significant[index] == 1){
      print(tmp_levels[index])
      rows_to_edit = which(data$genre == as.character(tmp_levels[index]))
      print(rows_to_edit)
      data$genre[rows_to_edit] = paste("(*)",tmp_levels[index])
    }
  }
  
  data
}
```

## Heatmap

```{r}
library(ggplot2)
library(RColorBrewer)
library(tidyr)
```

### Raw relative abundance

In this section, we will take the raw relative abundance.

First, we filter data by class and an abundance threshold.

```{r}
# Selected classes
selected_classes = c("Archaea", "Bacteria", "Eukaryota", "Viruses") 

filtered_data = subset(data, kingdom %in% selected_classes)
filtered_data = drop_na(filtered_data[c("kingdom","genre","ind","cat","abond")])

head(filtered_data)
```

Only those columns ar necessary in order to make the heatmap for the *genre* model.

```{r}
ggplot(filtered_data, aes(ind, genre, fill= abond)) + 
  geom_tile(colour = "white", size = 0.5)+
  scale_fill_distiller(palette = "Spectral", # Setting scale's palette
                       #limits = c(min(filtered_data$abond), # Setting scale limits
                        #          max(filtered_data$abond)),
                       direction = -1)+ # Setting direction of the palette
  scale_y_discrete(expand=c(0, 0))+ # Removing blank space between data and Y axis
  scale_x_discrete(expand=c(0, 0))+ # Same but for X axis
  labs(fill="Relative abundance (%)", # Setting legend and title name
       title = "Genre relative abundance across samples")+
  theme_grey(base_size=12)+ # Setting the base font size
  theme(line = element_blank(), # Removing the lines in between the boxes
        axis.ticks.x = element_blank(), # Removing ticks on X axis
        #axis.text.x = element_blank(), # Removing X labels
        axis.text.y = element_text(face="bold.italic"), # Setting Y labels to bold
        axis.title.x = element_blank(), # Removing X axis title
        axis.title.y = element_blank(), # Removing Y axis title
        panel.border=element_blank(), # Removing space between figure and panel
        legend.text=element_text(face="bold", size = 15), # Customizing legend font
        plot.title = element_text(face="bold", size = 25), # Same with main title
        legend.title = element_text(face="bold", size = 15), # Same with legend
        legend.key.size = unit(2, "cm"), # Setting the legend scale
        strip.text.y = element_text(angle = 0, face = "bold", size = 15),
        strip.text.x = element_text(face = "bold", size = 15))+
  facet_grid(rows = vars(filtered_data$kingdom), 
             cols = vars(factor(filtered_data$cat, 
                                levels = c("Control", "Colistin"))),
             scales = "free",
             space = "free")
```

While this looks awfull in this notebook, the exported image is fine. We then export the image :

```{r}
scale = 200 # Set the scale for the exported image

ggsave(filename = "../export/HM_taxonomy.png",
       width = 16*scale,
       height = 18*scale,
       units = "px",
       dpi=200)
```

### Log of the relative abundance

In this section, we will take the log of the relative abundance in order to see more subtile difference across samples.

```{r}
# Selected classes
selected_classes = c("Archaea", "Bacteria", "Eukaryota", "Viruses") 

filtered_data = subset(data, kingdom %in% selected_classes)
filtered_data = drop_na(filtered_data[c("kingdom","genre","ind","cat","abond")])

filtered_data$abond = log(filtered_data$abond)

head(filtered_data)
```

```{r}
ggplot(filtered_data, aes(ind, genre, fill= abond)) + 
  geom_tile(colour = "white", size = 0.5)+
  scale_fill_distiller(palette = "Spectral", # Setting scale's palette
                       #limits = c(min(filtered_data$abond), # Setting scale limits
                        #          max(filtered_data$abond)),
                       direction = -1)+ # Setting direction of the palette
  scale_y_discrete(expand=c(0, 0))+ # Removing blank space between data and Y axis
  scale_x_discrete(expand=c(0, 0))+ # Same but for X axis
  labs(fill="Log of the relative abundance", # Setting legend and title name
       title = "Genre relative abundance across samples")+
  theme_grey(base_size=12)+ # Setting the base font size
  theme(line = element_blank(), # Removing the lines in between the boxes
        axis.ticks.x = element_blank(), # Removing ticks on X axis
        #axis.text.x = element_blank(), # Removing X labels
        axis.text.y = element_text(face="bold.italic"), # Setting Y labels to bold
        axis.title.x = element_blank(), # Removing X axis title
        axis.title.y = element_blank(), # Removing Y axis title
        panel.border=element_blank(), # Removing space between figure and panel
        legend.text=element_text(face="bold", size = 15), # Customizing legend font
        plot.title = element_text(face="bold", size = 25), # Same with main title
        legend.title = element_text(face="bold", size = 15), # Same with legend
        legend.key.size = unit(2, "cm"), # Setting the legend scale
        strip.text.y = element_text(angle = 0, face = "bold", size = 15),
        strip.text.x = element_text(face = "bold", size = 15))+
  facet_grid(rows = vars(filtered_data$kingdom), 
             cols = vars(factor(filtered_data$cat, 
                                levels = c("Control", "Colistin"))),
             scales = "free",
             space = "free")
```

Finally, we export the graph.

```{r}
scale = 200 # Set the scale for the exported image

ggsave(filename = "../export/HM_taxonomy_(log).png",
       width = 16*scale,
       height = 18*scale,
       units = "px",
       dpi=200)
```

### Normalized version

First we need to normalize the data

```{r}
# Selected classes
selected_classes = c("Archaea", "Bacteria", "Eukaryota", "Viruses") 

filtered_data = subset(data, kingdom %in% selected_classes)
filtered_data = drop_na(filtered_data[c("kingdom","genre","ind","cat","abond")])

library(caret)
process = preProcess(as.data.frame(filtered_data$abond), method=c("range"))
filtered_data$abond = predict(process, as.data.frame(filtered_data$abond))[,1]

#filtered_data$abond = scale(filtered_data$abond)

head(filtered_data)
```

```{r}
ggplot(filtered_data, aes(ind, genre, fill= abond)) + 
  geom_tile(colour = "white", size = 0.5)+
  scale_fill_distiller(palette = "Spectral", # Setting scale's palette
                       #limits = c(min(filtered_data$abond), # Setting scale limits
                        #          max(filtered_data$abond)),
                       direction = -1)+ # Setting direction of the palette
  scale_y_discrete(expand=c(0, 0))+ # Removing blank space between data and Y axis
  scale_x_discrete(expand=c(0, 0))+ # Same but for X axis
  labs(fill="Normalized relative abundance", # Setting legend and title name
       title = "Genre relative abundance across samples")+
  theme_grey(base_size=12)+ # Setting the base font size
  theme(line = element_blank(), # Removing the lines in between the boxes
        axis.ticks.x = element_blank(), # Removing ticks on X axis
        #axis.text.x = element_blank(), # Removing X labels
        axis.text.y = element_text(face="bold.italic"), # Setting Y labels to bold
        axis.title.x = element_blank(), # Removing X axis title
        axis.title.y = element_blank(), # Removing Y axis title
        panel.border=element_blank(), # Removing space between figure and panel
        legend.text=element_text(face="bold", size = 15), # Customizing legend font
        plot.title = element_text(face="bold", size = 25), # Same with main title
        legend.title = element_text(face="bold", size = 15), # Same with legend
        legend.key.size = unit(2, "cm"), # Setting the legend scale
        strip.text.y = element_text(angle = 0, face = "bold", size = 15),
        strip.text.x = element_text(face = "bold", size = 15))+
  facet_grid(rows = vars(filtered_data$kingdom), 
             cols = vars(factor(filtered_data$cat, 
                                levels = c("Control", "Colistin"))),
             scales = "free",
             space = "free")
```

```{r}
scale = 200 # Set the scale for the exported image

ggsave(filename = "../export/HM_taxonomy_(normalized).png",
       width = 16*scale,
       height = 18*scale,
       units = "px",
       dpi=200)
```

## Boxplot

### Raw relative abundance

```{r}
# Selected classes
selected_classes = c("Archaea", "Bacteria", "Eukaryota", "Viruses") 

filtered_data = subset(data, kingdom %in% selected_classes)
filtered_data = drop_na(filtered_data[c("kingdom","genre","ind","cat","abond")])

#filtered_data = droplevels(filtered_data[order(filtered_data$abond, 
#                                               decreasing = T),])
#filtered_data$genre = factor(filtered_data$genre, levels = unique(filtered_data$genre))

#filtered_data = diff_test(filtered_data)

head(filtered_data)
```

```{r}
ggplot(filtered_data, aes(abond, genre, fill = cat)) + 
  geom_boxplot()+
  scale_y_discrete(expand=c(0, 0))+
  #scale_x_discrete(expand=c(0, 0))+
  labs(fill="Groups", title = "Relative abundance across samples")+
  xlab("Relative abundance (%)")+
  theme_grey(base_size=12)+
  theme(line = element_blank(),
        #axis.ticks.x = element_blank(),
        axis.text.x = element_text(face="bold"),
        axis.text.y = element_text(face="bold.italic"),
        #axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.border=element_blank(),
        legend.text=element_text(face="bold", size = 15),
        plot.title = element_text(face="bold", size = 25),
        legend.title = element_text(face="bold", size = 15),
        legend.key.size = unit(2, "cm"),
        strip.text.y = element_text(angle = 0, face = "bold", size = 15),
        strip.text.x = element_text(face = "bold", size = 15))+
  facet_grid(rows = vars(filtered_data$kingdom), 
             #cols = vars(factor(data$cat, levels = c("Control", "Colistin"))),
             scales = "free",
             space = "free")
```

And once again we export the figure.

```{r}
scale = 200
ggsave(filename = "../export/BP_taxonomy.png",
       width = 16*scale,
       height = 18*scale,
       units = "px",
       dpi=200)
```

### Log of the relative abundance

```{r}
# Selected classes
selected_classes = c("Archaea", "Bacteria", "Eukaryota", "Viruses") 

filtered_data = subset(data, kingdom %in% selected_classes)
filtered_data = drop_na(filtered_data[c("kingdom","genre","ind","cat","abond")])

#filtered_data = droplevels(filtered_data[order(filtered_data$abond, 
#                                               decreasing = T),])
#filtered_data$genre = factor(filtered_data$genre, levels = unique(filtered_data$genre))

#filtered_data = diff_test(filtered_data)

filtered_data$abond = log(filtered_data$abond)

head(filtered_data)
```

We generate the figure.

```{r}
ggplot(filtered_data, aes(abond, genre, fill = cat)) + 
  geom_boxplot()+
  scale_y_discrete(expand=c(0, 0))+
  #scale_x_discrete(expand=c(0, 0))+
  labs(fill="Groups", title = "Relative abundance across samples")+
  xlab("Log of the relative abundance")+
  theme_grey(base_size=12)+
  theme(line = element_blank(),
        #axis.ticks.x = element_blank(),
        axis.text.x = element_text(face="bold"),
        axis.text.y = element_text(face="bold.italic"),
        #axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.border=element_blank(),
        legend.text=element_text(face="bold", size = 15),
        plot.title = element_text(face="bold", size = 25),
        legend.title = element_text(face="bold", size = 15),
        legend.key.size = unit(2, "cm"),
        strip.text.y = element_text(angle = 0, face = "bold", size = 15),
        strip.text.x = element_text(face = "bold", size = 15))+
  facet_grid(rows = vars(filtered_data$kingdom), 
             #cols = vars(factor(data$cat, levels = c("Control", "Colistin"))),
             scales = "free",
             space = "free")
```

```{r}
scale = 200
ggsave(filename = "../export/BP_taxonomy_(log).png",
       width = 16*scale,
       height = 18*scale,
       units = "px",
       dpi=200)
```

### Normalized version

```{r}
# Selected classes
selected_classes = c("Archaea", "Bacteria", "Eukaryota", "Viruses") 

filtered_data = subset(data, kingdom %in% selected_classes)
filtered_data = drop_na(filtered_data[c("kingdom","genre","ind","cat","abond")])

#filtered_data = droplevels(filtered_data[order(filtered_data$abond, 
#                                               decreasing = T),])
#filtered_data$genre = factor(filtered_data$genre, levels = unique(filtered_data$genre))

#filtered_data = diff_test(filtered_data)

library(caret)
process = preProcess(as.data.frame(filtered_data$abond), method=c("range"))
filtered_data$abond = predict(process, as.data.frame(filtered_data$abond))[,1]

head(filtered_data)
```

```{r}
ggplot(filtered_data, aes(abond, genre, fill = cat)) + 
  geom_boxplot()+
  scale_y_discrete(expand=c(0, 0))+
  #scale_x_discrete(expand=c(0, 0))+
  labs(fill="Groups", title = "Relative abundance across samples")+
  xlab("Normalized relative abundance (%)")+
  theme_grey(base_size=12)+
  theme(line = element_blank(),
        #axis.ticks.x = element_blank(),
        axis.text.x = element_text(face="bold"),
        axis.text.y = element_text(face="bold.italic"),
        #axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.border=element_blank(),
        legend.text=element_text(face="bold", size = 15),
        plot.title = element_text(face="bold", size = 25),
        legend.title = element_text(face="bold", size = 15),
        legend.key.size = unit(2, "cm"),
        strip.text.y = element_text(angle = 0, face = "bold", size = 15),
        strip.text.x = element_text(face = "bold", size = 15))+
  facet_grid(rows = vars(filtered_data$kingdom), 
             #cols = vars(factor(data$cat, levels = c("Control", "Colistin"))),
             scales = "free",
             space = "free")
```

```{r}
scale = 200
ggsave(filename = "../export/BP_taxonomy_(normalized).png",
       width = 16*scale,
       height = 18*scale,
       units = "px",
       dpi=200)
```
